---
title: "Code for linear tests"
author: "D. Rex Mitchell"
date: '2022-09-14'
output: html_document
---
```{r}
library(rstudioapi) # Safely Access the RStudio API
setwd(dirname(getActiveDocumentContext()$path))
```

Load packages and import data
```{r}
library(geomorph)
library(nlme)
library(landvR)
data <- read.morphologika("../Data/Mrufus_morphologika.txt") # doesn't work in Rmd for some reason
```


### Heatmap function ###
```{r}
map2color <- function(x, pal, limits = range(x)){
  pal[findInterval(x, seq(limits[1], limits[2], length.out = length(pal) + 1), 
                   all.inside=TRUE)]
}
```

Procrustes superimpositions
```{r}
ruf_gpa <- gpagen(data$coords)
pairs <- read.csv("../Data/landpairs.csv", header=T) # not working in Rmd for some reason
ind <- dimnames(data$coords)[[3]]
ruf_gpa_bi<- bilat.symmetry(data$coords, ind = ind, land.pairs = pairs, object.sym=T)
summary(ruf_gpa_bi)
```

Extract variables
```{r}
mn <- mshape(ruf_gpa_bi$symm.shape)
shape <- ruf_gpa_bi$symm.shape
sex <- as.factor(data$labels[,1])  # 98F, 68M
pop <- as.factor(data$labels[,6])  # 115M, 51Q
size <- ruf_gpa$Csize
age <- data$labels[,5]
foot <- data$labels[,3]
wt <- data$labels[,2]
```

Chi square test of sex ratios
```{r}
nmat <- matrix(data=c(67.83,32.17,39.22,60.79), nrow=2, ncol=2, byrow=T, dimnames = list(c("M", "Q"),c("F", "M")))
cst <- chisq.test(nmat);cst
# sex ratios significantly different.
```

Check outliers
```{r}
plotOutliers(ruf_gpa$coords)
# several outliers, but no landmark anomalies. All are juveniles/subadults
```

Cranial shape allometry
```{r}
allom_all <- procD.lm(shape~log(size), iter=999); summary(allom_all)
# Given a particular size, sex has a very small influence on shape (Rsq = 0.010, p = 0.006)

allom_plot <- plot(allom_all, type = "regression", predictor = log(size), 
                    reg.type = "RegScore", pch = 21, bg="grey",
                    xlab = "log(centroid size)", ylab = "Regression score", cex=2)
line <- lm(allom_plot$RegScore~log(size)); line
abline(-1.3259,0.2332, lwd=3, lty=2)

preds <- shape.predictor(allom_all$GM$fitted, x = allom_plot$PredLine, 
                         predmin = min(allom_plot$PredLine), 
                         predmax = max(allom_plot$PredLine))
differences <- coordinates.difference(coordinates = preds$predmax,
                                      reference = preds$predmin, type = "spherical")
procrustes.var.plot(preds$predmin, preds$predmax, col = heat.colors, col.val = differences[[1]][1:37], magnitude = 1, pt.size = 1)
```

Sexual dimorphism in cranial shape
```{r}
allom_all <- procD.lm(shape~log(ruf_gpa$Csize)*as.factor(sex), iter=999); summary(allom_all)
# sex has a very small influence on shape (Rsq = 0.010, p = 0.006)

allom_plot <- plot(allom_all, type = "regression", predictor = log(size), reg.type = "RegScore", pch = c(21,24)[sex], bg=c("gray29","grey69")[sex], xlab = "log(centroid size)", ylab = "Regression score", cex=3)
```


### compare populations at same age ranges ###
First, subset data ranges
```{r}
plot(age~pop)

mage <- subset(age, pop=="M")
mage <- as.numeric(mage)
min(mage) # 1.149230479
max(mage) # 17.58357

qage <- subset(age, pop=="Q")
qage <- as.numeric(qage)
min(qage) # 2.520328
max(qage) # 14.90415

age <- as.numeric(age)
same_age_data <- subset(two.d.array(data$coords), age >= 2.520328 & age <= 14.90415)
sa_array <- arrayspecs(same_age_data, 37, 3)
names <- subset(dimnames(shape)[[3]], age > 2.520328 & age < 14.90415)
ind <- names
ruf_gpa_bi_sa<- bilat.symmetry(sa_array, ind = ind, land.pairs = pairs, object.sym=T)
shape_sa <- ruf_gpa_bi_sa$symm.shape
```

subset variables
```{r}
ps <- gpagen(sa_array)
size_sa <- ps$Csize
size_sa <- as.numeric(size_sa)
sex_sa <- subset(sex, age >= 2.520328 & age <= 14.90415) 
pop_sa <- subset(pop, age >= 2.520328 & age <= 14.90415)
foot_sa <- subset(foot, age >= 2.520328 & age <= 14.90415)
wt_sa <-subset(wt, age >= 2.520328 & age <= 14.90415)
age_sa <- subset(age, age >= 2.520328 & age <= 14.90415)
```


### split into sexes
```{r}
cran_saF <- subset(two.d.array(shape_sa), sex_sa=="F")
shape_saF <- arrayspecs(cran_saF, 37,3) # 86
cran_saM <- subset(two.d.array(shape_sa), sex_sa=="M")
shape_saM <- arrayspecs(cran_saM, 37,3) # 63
foot_saF <- subset(foot_sa, sex_sa=="F")
foot_saM <- subset(foot_sa, sex_sa=="M")
size_saF <- subset(size_sa, sex_sa=="F")
size_saM <- subset(size_sa, sex_sa=="M")
pop_saF <- subset(pop_sa, sex_sa=="F")
pop_saM <- subset(pop_sa, sex_sa=="M")
wt_saF <- subset(wt_sa, sex_sa=="F")
wt_saM <- subset(wt_sa, sex_sa=="M")
age_saF <- subset(age_sa, sex_sa=="F")
age_saM <- subset(age_sa, sex_sa=="M")
```

shape/age regressions
```{r}
Fageshape <- procD.lm(shape_saF~age_saF*pop_saF, iter=999); summary(Fageshape) # not sig
Mageshape <- procD.lm(shape_saM~age_saM*pop_saM, iter=999); summary(Mageshape) # not sig
```

Centroid size/age resressions
```{r}
Fagepop <- procD.lm(log(size_saF)~age_saF*pop_saF, iter=999); summary(Fagepop) # not sig
Magepop <- procD.lm(log(size_saM)~age_saM*pop_saM, iter=999); summary(Magepop) # sig population diffs (p = 0.006)

Magepop_plot <- plot(age_saM, log(size_saM), pch = c(21,24)[pop_saM], bg=c("orange","purple")[pop_saM], 
                     xlab = "age", ylab = "log(cranial size)", cex=3, cex.lab=1.5)

size_mm <- subset(size_saM, pop_saM=="M")
size_qm <- subset(size_saM, pop_saM=="Q")
age_mm <- subset(age_saM, pop_saM=="M")
age_qm <- subset(age_saM, pop_saM=="Q")

mline <- lm(log(size_mm)~age_mm) 
abline(5.58717, 0.03048, col = "orange", lwd=4, lty=3)
qline <- lm(log(size_qm)~age_qm)
abline( 5.65375,  0.02172, col = "purple", lwd=4, lty=5)
```

Weight/age regressions
```{r}
Fagepopwt <- procD.lm(wt_saF~age_saF*pop_saF, iter=999); summary(Fagepopwt)
Fagepopwt_plot <- plot(age_saF, wt_saF, pch = c(21,24)[pop_saF], bg=c("orange","purple")[pop_saF], 
                     xlab = "age", ylab = "body weight (kg)", cex=3, cex.lab=1.5)

wt_mf <- subset(wt_saF, pop_saF=="M")
wt_qf <- subset(wt_saF, pop_saF=="Q")
age_mf <- subset(age_saF, pop_saF=="M")
age_qf <- subset(age_saF, pop_saF=="Q")

mline <- lm(wt_mf~age_mf); mline 
abline(16.9108, 0.5617, col = "orange", lwd=4, lty=3)
qline <- lm(wt_qf~age_qf); qline
abline(18.3938, 0.6405, col = "purple", lwd=4, lty=5)

Magepopwt <- procD.lm(wt_saM~age_saM*pop_saM, iter=999); summary(Magepopwt)
Magepopwt_plot <- plot(age_saM, wt_saM, pch = c(21,24)[pop_saM], bg=c("orange","purple")[pop_saM], 
                       xlab = "age", ylab = "body weight (kg)", cex=3, cex.lab=1.5)

wt_mm <- subset(wt_saM, pop_saM=="M")
wt_qm <- subset(wt_saM, pop_saM=="Q")
age_mm <- subset(age_saM, pop_saM=="M")
age_qm <- subset(age_saM, pop_saM=="Q")

mline <- lm(wt_mm~age_mm); mline 
abline(11.711, 3.116, col = "orange", lwd=4, lty=3)
qline <- lm(wt_qm~age_qm); qline
abline(20.755, 1.934, col = "purple", lwd=4, lty=5)
```

Pes length/age regressions
```{r}
Fagepopfoot <- procD.lm(foot_saF~age_saF*pop_saF, iter=999); summary(Fagepopfoot)
Fagepopfoot_plot <- plot(age_saF, foot_saF, pch = c(21,24)[pop_saF], bg=c("orange","purple")[pop_saF], 
                       xlab = "age", ylab = "pes length (mm)", cex=3, cex.lab=1.5)

foot_mf <- subset(foot_saF, pop_saF=="M")
foot_qf <- subset(foot_saF, pop_saF=="Q")
age_mf <- subset(age_saF, pop_saF=="M")
age_qf <- subset(age_saF, pop_saF=="Q")

mline <- lm(foot_mf~age_mf); mline 
abline(274.461, 1.623, col = "orange", lwd=4, lty=3)
qline <- lm(foot_qf~age_qf); qline
abline(285.1298, 0.8303, col = "purple", lwd=4, lty=5)


Magepopfoot <- procD.lm(foot_saM~age_saM*pop_saM, iter=999); summary(Magepopfoot)
Magepopfoot_plot <- plot(age_saM, foot_saM, pch = c(21,24)[pop_saM], bg=c("orange","purple")[pop_saM], 
                         xlab = "age", ylab = "pes length (mm)", cex=3, cex.lab=1.5)

foot_mm <- subset(foot_saM, pop_saM=="M")
foot_qm <- subset(foot_saM, pop_saM=="Q")
age_mm <- subset(age_saM, pop_saM=="M")
age_qm <- subset(age_saM, pop_saM=="Q")

mline <- lm(foot_mm~age_mm); mline 
abline(280.373, 7.862, col = "orange", lwd=4, lty=3)
qline <- lm(foot_qm~age_qm); qline
abline(311.447, 3.081, col = "purple", lwd=4, lty=5)
```




